---
title: "Plotting field maps with the desplot package"
author: "Kevin Wright"
date: "`r Sys.Date()`"
bibliography: desplot.bib
output:
  html_document:
    highlight: tango
vignette: >
  % \VignetteIndexEntry{Plotting field maps with the desplot package}
  % \VignetteEngine{knitr::rmarkdown}
---
<style type="text/css">
body, td { font-size: 14pt; color: #000; padding:1em; }
code, pre, #ws, #message { font-size: 12pt;
  color: #000; background-color: #f5f5f5; }
code {border: 1px solid #e0e0e0; padding: 0 5px;}
pre > code { border: 0; }
</style>

# Abstract

This short note shows how to plot a field map from an agricultural experiment
and why that may be useful.

# R setup

```{r setup}
library("knitr")
knitr::opts_chunk$set(fig.align="center", fig.width=6, fig.height=6)
options(width=90)
```

## Example 1

First, a plot of the experimental design of the oats data from @yates1935complex.
```{r yates}
library(agridat)
library(desplot)
data(yates.oats)
# Older versions of agridat used x/y here instead of col/row
if(is.element("x",names(yates.oats)))
   yates.oats <- transform(yates.oats, col=x, row=y)
desplot(block ~ col+row, yates.oats,
        col=nitro, text=gen, cex=1, out1=block,
        out2=gen, out2.gpar=list(col = "gray50", lwd = 1, lty = 1))
```

## Example 2

This next example is from @ryder1981field.  Fit an ordinary RCB model with fixed
effects for `block` and `genotype`.  Plot a heatmap of the residuals.
```{r ryder}
library(agridat)
library(desplot)
data(ryder.groundnut)
gnut <- ryder.groundnut
m1 <- lm(dry ~ block + gen, gnut) # Standard RCB model
gnut$res <- resid(m1)
desplot(res ~ col + row, gnut, text=gen, cex=1,
        main="ryder.groundnut residuals from RCB model")
```
Note the largest positive/negative residuals are adjacent to each other,
perhaps caused by the original data values being swapped.  Checking with
experiment investigators (managers, data collectors, etc.) is recommended.

## Session info

```{r session}
sessionInfo()
```

## References
	
