---
title: "Plotting field maps with the desplot package"
author: "Kevin Wright"
date: "`r Sys.Date()`"
bibliography: desplot.bib
output:
  rmarkdown::html_vignette:
vignette: >
  %\VignetteIndexEntry{Plotting field maps with the desplot package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Abstract

This short note shows how to plot a field map from an agricultural experiment and why that may be useful.

# R setup

```{r setup}
library("knitr")
knitr::opts_chunk$set(fig.align="center", fig.width=6, fig.height=6)
options(width=90)
```

## Example 1

First, a plot of the experimental design of the oats data from @yates1935complex.
```{r yates}
library(agridat)
library(desplot)
data(yates.oats)
desplot(yates.oats, block ~ col+row,
        col=nitro, text=gen, cex=1, out1=block,
        out2=gen, out2.gpar=list(col = "gray50", lwd = 1, lty = 1))
```

## Example 2

This next example is from @ryder1981field.  Fit an ordinary RCB model with fixed effects for `block` and `genotype`.  Plot a heatmap of the residuals.

```{r ryder}
library(agridat)
library(desplot)
data(ryder.groundnut)
gnut <- ryder.groundnut
m1 <- lm(dry ~ block + gen, gnut) # Standard RCB model
gnut$res <- resid(m1)
desplot(gnut, res ~ col + row, text=gen, cex=1,
        main="ryder.groundnut residuals from RCB model")
```
Note the largest positive/negative residuals are adjacent to each other,
perhaps caused by the original data values being swapped.  Checking with
experiment investigators (managers, data collectors, etc.) is recommended.


## References
	
